```{r}
library(rvest)
library(httr)
library(nhdplusTools)
library(ggplot2)
library(sf)
library(dplyr)
library(readr)

```

Download in the data sets:

1.  Active SNOTEL: Active stations data
2.  \_\_\_\_\_\_\_\_\_\_\_\_\_\_: 1991-2020 Median SWE data
3.  \_\_\_\_\_\_\_\_\_\_\_\_\_\_: YTD SWE data
4.  \_\_\_\_\_\_\_\_\_\_\_\_\_\_: Watershed Map data
5.  \_\_\_\_\_\_\_\_\_\_\_\_\_\_: Map of USA data
6.  \_\_\_\_\_\_\_\_\_\_\_\_\_\_: Hydrograph YTD data

```{r}
##Data Wrangling: 

# Step 1: Pull URL
url <- "https://wcc.sc.egov.usda.gov/reports/UpdateReport.html?report=Western+US"
page <- read_html(url)

# Step 2: Pull table
tables <- html_nodes(page, "table")
df2 <- html_table(tables[[2]], fill = TRUE)
View(df2)
names(df2)[1] <- "name"

##Watershed Data Table (Assistance from AI to find this )
install.packages(c("nhdplusTools", "ggplot2", "sf", "dplyr"))
western_huc4 <- get_huc(
  id = c("1901","1902","1903","1904","1905","1906","1001","1002","1003","1004","1005","1006","1007","1008",
         "1301","1302","1303","1401","1402","1403","1404","1405","1406","1407",
         "1501","1502","1503","1504","1505","1506","1507","1508",
         "1601","1602","1603","1604",
         "1701","1702","1703","1704","1705","1706","1707","1708",
         "1801","1802","1803","1804","1805"),
  type = "huc04"
)


# Basic US state background Data
us_states <- map_data("state")
western <- us_states |>
  filter(region %in% c("oregon","california","nevada","arizona",
                        "wyoming","colorado","utah","montana","idaho",
                        "washington","new mexico", "alaska"))
## I want to make my data into CSV to make it easier so just incase my code breaks this is my safety
write_csv(df2, "PercentSWE.csv")
write_csv(western_huc4, "westernhuc4.csv")

##Skipping the first two columns cleans my data also renaming
PercentSWE <- read.csv("PercentSWE.csv", skip = 2, header = FALSE, stringsAsFactors = FALSE)
colnames(PercentSWE) <- c("name", "SWE_sites", "SWE_pct_median", "precip_sites", "precip_pct_median", "region")

##Recoding WesternHuc4
western_huc4 <- western_huc4 |>
  mutate(name = toupper(name))

##Matching all the names 
PercentSWE <- PercentSWE |>
  mutate(name = case_when(
    name == "GUNNISON RIVER BASIN"                        ~ "GUNNISON",
    name == "UPPER COLORADO RIVER BASIN"                  ~ "COLORADO HEADWATERS",
    name == "YAMPA AND WHITE RIVER BASINS"                ~ "WHITE-YAMPA",
    name == "DIRTY DEVIL"                                 ~ "UPPER COLORADO-DIRTY DEVIL",
    name == "UPPER RIO GRANDE BASIN"                      ~ "RIO GRANDE HEADWATERS",
    name == "UPPER RIO GRANDE BASIN - New Mexico"         ~ "RIO GRANDE-ELEPHANT BUTTE",
    name == "MIMBRES RIVER BASIN"                         ~ "RIO GRANDE-MIMBRES",
    name == "GILA RIVER BASIN"                            ~ "UPPER GILA",
    name == "SALT RIVER BASIN"                            ~ "SALT",
    name == "LITTLE COLORADO RIVER BASIN"                 ~ "LITTLE COLORADO",        
    name == "SMITH, JUDITH, AND MUSSELSHELL RIVER BASINS" ~ "MISSOURI-MUSSELSHELL",
    name == "SUN, TETON AND MARIAS RIVER BASINS"          ~ "MISSOURI-MARIAS",
    name == "MISSOURI MAINSTEM RIVER BASIN"               ~ "MISSOURI-POPLAR",
    name == "MILK RIVER BASIN"                            ~ "MILK",
    name == "UPPER YELLOWSTONE RIVER BASIN"               ~ "UPPER YELLOWSTONE",
    name == "BIGHORN RIVER BASIN (WYOMING)"               ~ "BIG HORN",
    name == "SAN JUAN RIVER BASIN"                        ~ "UPPER COLORADO-DOLORES",
    name == "UPPER GREEN RIVER"                           ~ "GREAT DIVIDE-UPPER GREEN",
    name == "LOWER GREEN RIVER"                           ~ "LOWER GREEN",
    name == "LOWER SNAKE, WALLA WALLA"                    ~ "LOWER SNAKE",
    name == "UPPER HUMBOLDT RIVER"                        ~ "BLACK ROCK DESERT-HUMBOLDT",
    name == "BEAR RIVER"                                  ~ "BEAR",
    name == "ESCALANTE RIVER"                             ~ "ESCALANTE DESERT-SEVIER LAKE",
    name == "LOWER PEND OREILLE"                          ~ "KOOTENAI-PEND OREILLE-SPOKANE",
    name == "UPPER YAKIMA"                                ~ "YAKIMA",
    name == "CENTRAL COLUMBIA"                            ~ "MIDDLE COLUMBIA",
    name == "KLAMATH"                                     ~ "KLAMATH-NORTHERN CALIFORNIA COASTAL",
    name == "NORTHERN GREAT BASIN"                             ~ "GREAT SALT LAKE",        
    name == "SNAKE BASIN ABOVE PALISADES"                 ~ "UPPER SNAKE",            
    name == "SNAKE BASIN ABOVE AMERICAN FALLS"            ~ "MIDDLE SNAKE",
    name == "SOUTHEAST ALASKA BASIN"            ~ "SOUTHEAST ALASKA",
    name == "SOUTHEAST ALASKA BASIN"            ~ "SOUTH CENTRAL ALASKA",
    TRUE ~ name
  ))

##Remove * and ""

PercentSWE <- PercentSWE |>
  mutate(SWE_pct_median = as.numeric(gsub("\\*", "", SWE_pct_median)))

##Join 
df_joined <- western_huc4 |>
  left_join(PercentSWE, by = "name")


```
## Plot
```{r}
# Plot - fill by SWE percent of median, gray for unmatched basins
ggplot() +
  geom_polygon(data = western, aes(x = long, y = lat, group = group),
               fill = "gray90", color = "white") +
  geom_sf(data = df_joined, aes(fill = SWE_pct_median),
          color = "steelblue", linewidth = 0.6) +
  geom_sf_text(data = df_joined, aes(label = name),
               size = 1.8) +
  scale_fill_gradient2(low = "red", mid = "white", high = "blue",
                       midpoint = 100,
                       na.value = "gray80",
                       name = "% of Median SWE") +
  coord_sf(xlim = c(-170, -102), ylim = c(30, 72)) +
  theme_minimal() +
  labs(title = "Western US HUC4 Watershed Basins - % of Median SWE")

##Observe no values are over 100, which is the midpoint and the median from the last 30 years. 
```





Mutate and Join data:

1.  Combined \<- Median to YTD values to calculate percent
2.  Combined join with watershed map data
3.  Watershed map data join with USA map data

Plot:

1.  Create visualization of USA with %SWE per watershed
2.  Create Hydrograph Visualization

Shiny:

1.  Select State
2.  Select Watershed
3.  Display Hydrograph YTD plot
4.  Return a message detailing if drought is likely
